---
title: "07_epi_abm"
output: html_document
---

# ODEs of the bi_strain model 

The equations below are describing a compartmental model 

\begin{align*}
\frac{dS_p}{dt} &= (1 - \lambda_{cr} - \lambda_{cs} - \lambda_{ir} - \lambda_{is})\Lambda - (\beta_{np}^r + \beta_{np}^s)S_p - \tau_s S_p \\
\frac{dC_p^s}{dt} &= \lambda_{cs}\Lambda + (1-x)\beta_{np}^s S_p - (\gamma_s + \tau_s + \mu_c)C_p^s \\  
\frac{dI_p^s}{dt} &= \lambda_{is}\Lambda + x\beta_{np}^s S_p + \gamma_s C_p^s - (\Psi_s + \mu_i^s)I_p^s \\  
\frac{dC_p^r}{dt} &= \lambda_{cr}\Lambda + (1-x)\beta_{np}^r S_p - (\gamma_r + \tau_r + \mu_c)C_p^r \\  
\frac{dI_p^r}{dt} &= \lambda_{ir}\Lambda + x\beta_{np}^r S_p + \gamma_r C_p^r - (\Psi_r + \mu_i^r)I_p^r \\  
\frac{dU_n}{dt} &= -(\beta_{pn}^r + \beta_{pn}^s) U_n + \delta (C_n^r + C_n^s) \\  
\frac{dC_n^s}{dt} &= \beta_{pn}^s U_n - \delta C_n^s \\  
\frac{dC_n^r}{dt} &= \beta_{pn}^r U_n - \delta C_n^r  
\end{align*}

To the best of my coding knowledge I have been able to come up with these custom modules below that more or less illustrates what happens in the above ODEs, but applied to an ICM using the EpiModel framework. 

below is the name of the different parameters involved in the ICM

```{r}

params <- list(
  psi_s    = NA,  # Treatment rate from MSSA infection
  psi_r    = NA,  # Treatment rate from MRSA infection
  iota     = NA,  # Treatment success probability
  s_b      = NA,  # Fitness cost
  s_p      = NA,  # Penalty on treatment efficacy due to resistance
  a        = NA,  # Contact rate between patients and HCWs
  b_pn     = NA,  # Transmission probability from patient to HCW
  theta    = NA,  # Precaution compliance of HCWs
  n        = NA,  # Penalty on transmission capabilities in colonized patients
  low      = NA,  # Lower threshold for decontamination of HCW
  high     = NA,  # Upper threshold for decontamination of HCW
  kappa    = NA,  # Duration of colonization (inverse = clearance rate)
  m_s      = NA,  # Probability of developing infection when colonized
  delta_m  = NA,  # Increment in m_s when colonized by resistance strain
  x        = NA,  # Proportion of exposed who develop infection directly
  b_np     = NA   # Transmission probability from HCW to patient
)
```


```{r Custom initialize.FUN}

initialize_module <- function(dat, at) {
  
  # Get parameters
  initial_pop_size <- get_param(dat, "initial_pop_size")
  prob_patient <- get_param(dat, "prob_patient", 0.85)  # Default: 85% patients
  prob_status_patients <- get_param(dat, "prob_status_patients", 
                                   c(0.80, 0.07, 0.05, 0.05, 0.03))  # Sp, Cp_s, Cp_r, Ip_s, Ip_r

  # Initialize core attributes
  active <- rep(1, initial_pop_size)
  type <- sample(c("patient", "nurse"), 
                 initial_pop_size, 
                 replace = TRUE, 
                 prob = c(prob_patient, 1 - prob_patient))
  status <- rep(NA, initial_pop_size)

  # Assign patient statuses
  patients <- which(type == "patient")
  status[patients] <- sample(
    c("Sp", "Cp_s", "Cp_r", "Ip_s", "Ip_r"),
    length(patients),
    replace = TRUE,
    prob = prob_status_patients
  )

  # Nurses start uncontaminated
  status[which(type == "nurse")] <- "Un"

  # Set attributes
  dat <- set_attr(dat, "active", active)
  dat <- set_attr(dat, "type", type)
  dat <- set_attr(dat, "status", status)

  # Track initial counts (for summary stats)
  dat <- set_epi(dat, "init.num.Sp", at, sum(status == "Sp"))
  dat <- set_epi(dat, "init.num.Cp_s", at, sum(status == "Cp_s"))
  dat <- set_epi(dat, "init.num.Cp_r", at, sum(status == "Cp_r"))
  dat <- set_epi(dat, "init.num.Ip_s", at, sum(status == "Ip_s"))
  dat <- set_epi(dat, "init.num.Ip_r", at, sum(status == "Ip_r"))
  dat <- set_epi(dat, "init.num.Un", at, sum(status == "Un"))

  return(dat)
  
}
```


```{r Module 1}

infection_module <-function(dat, at) {
  
  #Get attributes
  active <- get_attr(dat, "active")
  status <- get_attr(dat, "status")
  type <- get_attr(dat, "type")
  
  #Initialize trackers
  nInf_s <- 0
  nInf_r <- 0
  nCol_s <- 0
  nCol_r <- 0
  
  #Find susceptible patients
  idEligs <- which(active == 1 & type == "patient" & status == "Sp")
  
  if(length(idEligs >0)){
    
    #Call parameters
    a <- get_param(dat, "a")
    x <- get_param(dat, "x")
    s_b <- get_param(dat, "s_b")
    b_np <- get_param(dat, "b_np")
    theta <- get_param(dat, "theta")
    
    #what is the contamination pressure
    Cn_s <- sum(active == 1 & type == "nurse" & status == "Cn_s")
    Cn_r <- sum(active == 1 & type == "nurse" & status == "Cn_r")
    Nn <- sum(active == 1 & type == "nurse")
    
    #Define the force of infection
    beta_np_s <- a * b_np * (1-theta) * Cn_s/Nn
    beta_np_r <- a * b_np * (1 - theta) * (1 - s_b) * Cn_r / Nn
    
    #Derive prob from FOI
    probInf_s <- 1 - exp(-beta_np_s)
    probInf_r <- 1 - exp(-beta_np_r)
    
    #Simulate infections
    inf_s <- rbinom(length(idEligs), 1, probInf_s)
    inf_r <- rbinom(length(idEligs), 1, probInf_r)
    
    #Setting initial infection processes
    for(i in seq_along(idEligs)){
      
      id <- idEligs[i]
      
      if(inf_s[i] == 1 && inf_r[i] == 0){
        
        if(runif(1) < x){
          status[id] <- "Ip_s"
          nInf_s <- nInf_s + 1
        } else {
          status[id] <- "Cp_s"
          nCol_s <- nCol_s + 1
        }
        
      } else 
        
      if(inf_s[i] == 0 && inf_r[i] == 1){
        
        if(runif(1) < x){
          status[id] <- "Ip_r"
          nInf_r <- nInf_r + 1
        } else {
          status[id] <- "Ip_r"
          nCol_r <- nCol_r + 1
        }
      }
    }
    
  # Update attributes
  dat <- set_attr(dat, "status", status)  
    
  }
  
  # Record flows
  dat <- set_epi(dat, "np.inf.flow.s", at, nInf_s)
  dat <- set_epi(dat, "np.inf.flow.r", at, nInf_r)
  
  return(dat) 
  
}

```

```{r Module 2}

progression_module <- function(dat, at) {
  
  # Get attributes
  active <- get_attr(dat, "active")
  status <- get_attr(dat, "status")
  
  # Initialize trackers
  nProg_s <- 0
  nProg_r <- 0
  
  # Find colonized patients
  idsElig_s <- which(active == 1 & status == "Cp_s")
  idsElig_r <- which(active == 1 & status == "Cp_r")
  
  # Get parameters
  kappa <- get_param(dat, "kappa")
  m_s <- get_param(dat, "m_s")
  delta_m <- get_param(dat, "delta_m")
  
  # Calculate progression rates
  gamma_s <- kappa * m_s
  gamma_r <- kappa * m_s * (1 + delta_m)
  
  # Simulate progression
  if (length(idsElig_s) > 0) {
    
    prog_s <- rbinom(length(idsElig_s), 1, gamma_s)
    nProg_s <- sum(prog_s)
    
    if (nProg_s > 0) {
      
      status[idsElig_s[which(prog_s == 1)]] <- "Ip_s"
    }
  }
  
  if (length(idsElig_r) > 0) {
    
    prog_r <- rbinom(length(idsElig_r), 1, gamma_r)
    nProg_r <- sum(prog_r)
    
    if (nProg_r > 0) {
    
      status[idsElig_r[which(prog_r == 1)]] <- "Ip_r"
    }
  }
  
  # Update attributes
  dat <- set_attr(dat, "status", status)
  
  # Record flows
  dat <- set_epi(dat, "prog.flow.s", at, nProg_s)
  dat <- set_epi(dat, "prog.flow.r", at, nProg_r)
  
  return(dat)
  
}
```

```{r Module 3}

treatment_module <- function(dat, at) {
  
  # Get attributes
  active <- get_attr(dat, "active")
  status <- get_attr(dat, "status")
  
  # Initialize trackers
  nTreat_s <- 0 
  nTreat_r <- 0
  nRecov_s <- 0
  nRecov_r <- 0
  nDischarge <- 0
  
  # Find infected patients
  idsElig_s <- which(active == 1 & status == "Ip_s")
  idsElig_r <- which(active == 1 & status == "Ip_r")
  
  # Get parameters
  psi_s <- get_param(dat, "psi_s")
  psi_r <- get_param(dat, "psi_r")
  iota <- get_param(dat, "iota")
  s_b <- get_param(dat, "s_b")
  s_p <- get_param(dat, "s_p")
  
  # MSSA treatment
  if (length(idsElig_s) > 0) {
    
    treat_s <- rbinom(length(idsElig_s), 1, psi_s)
    nTreat_s <- sum(treat_s)
    
    if (nTreat_s > 0) {
      treated_s <- idsElig_s[which(treat_s == 1)]
      recover <- rbinom(nTreat_s, 1, iota)
      nRecov_s <- sum(recover)
      status[treated_s[which(recover == 1)]] <- "Sp"
      active[treated_s[which(recover == 0)]] <- 0
      nDischarge <- nDischarge + sum(recover == 0)
    }
  }
  
  # MRSA treatment
  if (length(idsElig_r) > 0) {
    
    treat_r <- rbinom(length(idsElig_r), 1, psi_r)
    nTreat_r <- sum(treat_r)
    
    if (nTreat_r > 0) {
      treated_r <- idsElig_r[which(treat_r == 1)]
      recover <- rbinom(nTreat_r, 1, iota * (1 - s_p))
      nRecov_r <- sum(recover)
      status[treated_r[which(recover == 1)]] <- "Sp"
      active[treated_r[which(recover == 0)]] <- 0
      nDischarge <- nDischarge + sum(recover == 0)
    }
  }
  
  # Update attributes
  dat <- set_attr(dat, "status", status)
  dat <- set_attr(dat, "active", active)
  
  # Record flows
  dat <- set_epi(dat, "treat.flow.s", at, nTreat_s)
  dat <- set_epi(dat, "treat.flow.r", at, nTreat_r)
  dat <- set_epi(dat, "recov.flow.s", at, nRecov_s)
  dat <- set_epi(dat, "recov.flow.r", at, nRecov_r)
  dat <- set_epi(dat, "discharge.flow", at, nDischarge)
  
  return(dat)
  
}
```

```{r Module 4}

nurse_contamination_module <- function(dat, at) {
  
  # Get attributes
  active <- get_attr(dat, "active")
  status <- get_attr(dat, "status")
  type <- get_attr(dat, "type")
  
  # Initialize trackers
  nContam_s <- 0
  nContam_r <- 0
  
  # Find uncontaminated nurses
  idsElig <- which(active == 1 & type == "nurse" & status == "Un")
  
  if (length(idsElig) > 0) {
    
    # Get parameters
    a <- get_param(dat, "a")
    b_pn <- get_param(dat, "b_pn")
    theta <- get_param(dat, "theta")
    s_b <- get_param(dat, "s_b")
    n <- get_param(dat, "n")
    
    # Calculate contamination pressures
    Np <- sum(active == 1 & type == "patient")
    phi <- Np / length(idsElig)
    Cp_s <- sum(active == 1 & status == "Cp_s")
    Ip_s <- sum(active == 1 & status == "Ip_s")
    Cp_r <- sum(active == 1 & status == "Cp_r")
    Ip_r <- sum(active == 1 & status == "Ip_r")
    
    beta_pn_s <- a * b_pn * (1 - theta) * phi * ((n * Cp_s + Ip_s) / Np)
    beta_pn_r <- a * b_pn * (1 - theta) * phi * (1 - s_b) * ((n * Cp_r + Ip_r) / Np)
    
    #Convert to probabilities
    probCtm_s <- 1 - exp(-beta_pn_s)
    probCtm_r <- 1 - exp(-beta_pn_r)
    
    # Simulate contamination
    contam_s <- rbinom(length(idsElig), 1, probCtm_s)
    contam_r <- rbinom(length(idsElig), 1, probCtm_r)
    
    # Process contamination
    for (i in seq_along(idsElig)) {
      
      id <- idsElig[i]
      
      if (contam_s[i] == 1 && contam_r[i] == 0) {
        
        status[id] <- "Cn_s"
        nContam_s <- nContam_s + 1
      } else 
        
      if (contam_r[i] == 1 && contam_s[i] == 0) {
        
        status[id] <- "Cn_r"
        nContam_r <- nContam_r + 1
      }
      
    }
    
    
    # Update attributes
    dat <- set_attr(dat, "status", status)
  }
  
  # Record flows
  dat <- set_epi(dat, "pn.contam.flow.s", at, nContam_s)
  dat <- set_epi(dat, "pn.contam.flow.r", at, nContam_r)
  
  return(dat)
}
```

```{r Module 5}

decontamination_module <- function(dat, at) {
  
  # Get attributes
  active <- get_attr(dat, "active")
  status <- get_attr(dat, "status")
  
  # Initialize trackers
  nDecon_s <- 0
  nDecon_r <- 0
  
  # Find contaminated nurses
  idsElig_s <- which(active == 1 & status == "Cn_s")
  idsElig_r <- which(active == 1 & status == "Cn_r")
  
  # Get parameters
  theta <- get_param(dat, "theta")
  low <- get_param(dat, "low")
  high <- get_param(dat, "high")
  
  # Calculate decontamination rate
  delta <- (1 - theta) * low + theta * high
  
  # Simulate decontamination
  if (length(idsElig_s) > 0) {
    
    decon_s <- rbinom(length(idsElig_s), 1, delta)
    nDecon_s <- sum(decon_s)
    
    if (nDecon_s > 0) {
      
      status[idsElig_s[which(decon_s == 1)]] <- "Un"
    }
  }
  
  if (length(idsElig_r) > 0) {
    
    decon_r <- rbinom(length(idsElig_r), 1, delta)
    nDecon_r <- sum(decon_r)
    
    if (nDecon_r > 0) {
      
      status[idsElig_r[which(decon_r == 1)]] <- "Un"
    }
  }
  
  # Update attributes
  dat <- set_attr(dat, "status", status)
  
  # Record flows
  dat <- set_epi(dat, "decon.flow.s", at, nDecon_s)
  dat <- set_epi(dat, "decon.flow.r", at, nDecon_r)
  
  return(dat)
  
}

```

